The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-02 paper.tex
changes in the structure of section 2: split into two sections

The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like
designs of objects and non-human agents can evoke both positive and negative feelings 
of affinity in the human observer / user. A positive valence reflects the experience of 
emotional engagement with and feelings of empathy for the non-human agent
\citep{cheetham_human_2011}. This in turn may lead to an effective (human-like and 
human-social) interaction with the non-human agent, and increase people's acceptance of
it. 
However, at a certain point as the non-human agent
approaches but fails to closely resemble a human, the evoked valence in the
human user suddenly turns into a negative uncanny feeling of what
\cite{mori_uncanny_1970} describes as disgust and revulsion. This, in turn,
might lead to rejection of the non-human agent. The Uncanny
Valley hypothesis predicts that with further increasing the human-likeness of
the non-human agent such that it appears so realistic that it is not
distinguishable from a human but perceived to be human, the effect is again a
positive one, reflecting feelings and engagement that are similar as those
evoked by a human-human interaction. 
Of course, developers and designers of artificial agents
(\eg virtual characters, robots) try to exploit the positive effects of 
\textbf{anthropomorphic form} and avoid ``falling'' into this Uncanny Valley.
Consequently, there is a considerable research effort to
better understand the boundaries of anthropomorphic design and their effects
on the interaction and experience. To date, findings have been however
inconsistent \citep{cheetham_human_2011}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		PART 1: WHAT OTHERS DID AND WE ALREADY KNOW ABOUT ANTHROPOMORPHISM
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Anthropomorphism is an interesting phenomenon because it tells us something about 
the relationship between the human who anthropomorphizes, and the non-human agent 
which is anthropomorphized. Though commonly observed, it still appears controversial, at 
the same time, that humans tend to \textit{``see human in the non-
human''} \citep{epley_seeing_2007}. 
To shed light on this phenomenon as a whole, first of all, we would like to propose a clarification of terms.

\paragraph{Terminology\\}

@@

However, the term anthropomorphism should \textit{not} be used to refer to the design of a non-human agent or artifact. This would violate the understanding of anthropomorphism arising in an interaction. We propose to use the term  \textbf{anthropomorphic design} to refer to an imitation of human-like form in the \textit{form} of artificial agents. (In the following, we might use the terms \textit{form} and \textit{design} interchangeably.)

In literature, this distinction is not always done, which is a major source of confusion.\\

@@

As anthropomorphism arises in an interaction between a human and a non-human agent
\citep{persson_anthropomorphism_2000}, one can also
take two main \textit{perspectives} when trying to explain why humans anthropomorphize \citep{lee_human_2005}: an artifact-centered perspective and a human-centered perspective (see grey triangles in Figure \ref{fig:anthropofig}).

We now first present an artifact-centered perspective, and outline how 
robotics uses \textit{anthropomorphic design} to encourage 
people to interact in a social, human-like way with robots.
Then, we present a human-centered 
perspective and shed light on \textit{anthropomorphism as a social phenomenon}. We present 
theories from developmental psychology, cognitive science, and social psychology to explain people's tendency to 
perceive human-like characteristics in non-human agents.
It has to be kept in mind that both perspectives are inseparable; they apply at the 
same time, and we would like to present them here just as two different viewpoints 
on the same phenomenon.
In the end of this section, we synthesize 
the presented backgrounds into an extended understanding of anthropomorphism 
as a multi-factor phenomenon.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		ARTIFACT-CENTERED PERSPECTIVE: ANTHROPOMORPHIC DESIGN
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{An artifact-centered perspective on anthropomorphism}
\label{sec:anthropomorphic-design}

An artifact-centered perspective explains anthropomorphism from the ascertainable 
parts of the non-human agent, namely the anthropomorphic form of it.
Using the term \textit{form} to describe the design of an artifact, we refer 
not only to the shape (how it looks like) but to the total 
expression of the artifact \citep{bartneck_shaping_2004}. \cite{disalvo_hug:_2003} 
argues that \textit{form} includes a product's physical shape, materials, and 
behavioral qualities. The behavioral aspect is of special importance for interactive 
systems that show some degree of autonomy (independence).
Several categories of anthropomorphic form can be classified, but there is no 
common agreement on the borders and transitions between these categories: 
anthropomorphic, zoomorphic, caricatured, functional \citep{fong_survey_2003}.
The use of nature-inspired and in particular anthropomorphic (human-like) form in the 
design of non-human things is as old and manifold as people's tendency to anthropomorphize. 
Anthropomorphic (and life-like) forms have a long-standing tradition in design in general, and in the design of daily life things, in particular 
\citep{norman_design_1988,guthrie_bottles_2000}. Interestingly, from its very
beginning robots were resembling a human being in their appearance. Starting
from when the term \textit{robot} first appeared in a play by Karel \^{C}apek in the 1920ies, robots were man-made beings, created as the \textit{``shadow of man''} \citep{replicant_????}.

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize it. 
This perspective assumes that humans directly (mindlessly) respond to life-like and social 
cues that the non-human agent or system emits \citep{nass_machines_2000}. Without thoughtful mental processing, 
humans tend to simply apply stereotypes and heuristics to the non-human agent, and in turn apply human-human social schemas and norms to the occurring interactions.
\cite{takayama_perspectives_2012} propose to refer to this immediate (sometimes visceral) sense in a situation as \textit{in-the-moment}, compared to a more distanced cogitation and consideration which she calls \textit{reflective}. (As this distinction implies different cognitive processes, we will come back to it later, in the human-centered perspective, Section \ref{sec:human-centered}).

The view of people mindlessly reacting in a human-social manner to anthropomorphic forms
is based on the fact that humans naturally react to specific 
cues. From early childhood 
on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
or to respond to social cues, such as facial expressions, social gestures or turn-taking. Humans have a propensity toward particular cues that seem to trigger automatic interpersonal social responses \citep{takayama_perspectives_2012}. 
One has to keep in mind that in 
human-human interaction these cues are of various senses and modalities, 
and consequently human-like design is also realized addressing several senses. 
One of the most outstanding aspects is probably the outer appearance of an agent or 
artifact (in particular when it is unknown or unfamiliar). \cite{schmitz_concepts_2011} 
argues that an agent's outer appearance is likely to have an important impact on how 
the object is overall perceived. Consequently, if a robot's outer form resembles a 
human, the robot is likely to be treated similar to a human. Besides the physical shape of a robot, other more subtle anthropomorphic forms can be found in robots, including human-social behavioral cues (\eg eye-gaze), social gestures (\eg waving), or verbal politeness (\eg greeting, apologizing).

People's natural reaction to these and other anthropomorphic design cues of
robots (in particular, social or also called \textit{sociable} robots 
\citep{breazeal_sociable_2000}) are extensively studied. This research is of particular interest in human-robot interaction as the consequences of anthropomorphic design can be both of positive and negative nature, and it is yet not clear what exactly it is that makes the difference.

\subsubsection{How people perceive and interact with human-like robots\\}
\label{sec:anthropomorphism-robotics}

Most work that studies how far human users treat robots as social agents, builds on the 
famous work of \cite{reeves_media_1996} in HCI. In the so-called \textit{Media 
Equation}, the authors describe that people mindlessly react in a social way to their computer 
and that they tend to even attribute human (social) characteristics to it -- despite 
the fact that a computer does not have human-like characteristics. Apart from 
computers, anthropomorphism has also been observed with digital / virtual characters 
\citep{von_der_putten_it_2010}, and artificial embodied agents, 
such as robots \citep{rosenthal-vonderputten_experimental_2013}. 
In short, one can say that with robots, and human-like robots in 
particular, anthropomorphism is a commonly observed phenomenon. 
But what are the motivations which underlie the efforts of enhancing anthropomorphism 
through specific anthropomorphic design in robots? Why are anthropomorphic forms so 
distinct and at the same time controversial in robotics? There is no one single answer 
to these questions.

The underlying central point about the human-like design of robots lies in the 
implications that this design holds, and the suggestions for interactions that
it makes to the human user. 
In other words, the appearance 
and function of a product, whether it is an appliance or a humanoid robot, affects how 
people perceive it, interact with it, and build long-term relationships with 
it \citep{bartneck_shaping_2004}. 
The principle of human-like form is probably most prominent in the design of robots, 
due to several reasons, such as to foster efficient interaction and increase 
acceptance. An anthropomorphic design of a robot encourages people to reason about this 
unfamiliar system \emph{as if} it was a human. To increase this tendency, robots are 
designed with anthropomorphic (or lifelike) forms, or are able to display and use 
human-social cues when interacting with humans.
However, anthropomorphic robot designs have been controversially discussed and received 
remarkable attention since the early 1970ies, related to Mori's
theory of the \textit{Uncanny Valley} \citep{mori_uncanny_1970}.
%see Figure \ref{fig:uncanny_valley}.

%\begin{figure}\centering
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[scale=0.75]{uncanny-valley.jpg}
%% figure caption is below the figure
% \caption{Mashiro Mori, illustration of the \textit{Uncanny Valley} hypothesis. The curve models the user's perceived affinity with the artifact, as its human likeness is increased. Up to a certain (not yet defined) point, a person perceives more affinity in a more human-like agent; however with an abrupt shift towards a negative perception as soon as the agent surpasses this point, of resembling a human too closely but not fully. Originally from 1970; picture from a manuscript version of \cite{mori_uncanny_2012}}
% \label{fig:uncanny_valley}       % Give a unique label
% \end{figure}

Despite the fact that the Uncanny Valley theory was more a working hypothesis and not 
meant to be empirically examined \citep{cheetham_human_2011}, it has been studied quite 
extensively. However, results are inconsistent. We try to give a short overview (not 
exclusive) on work related to HRI and robotics in the following.

\subsubsection{Anthropomorphic form can be advantageous for HRI\\}

On one hand, developers and designers have discovered that human-like robots may better 
match the affordances that this to humans adapted world brings along. 
Apart from this, it is suggested that an increased human-likeness of a robot (up to a 
certain threshold) will generally lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970}.
% (see Figure \ref{fig:uncanny_valley}).
This further increases anthropomorphism, and as such
appears to be beneficial for human-robot relations, especially with socially 
interactive robots \citep{fong_survey_2003}. Consequently, the domain of (social) 
robotics tries to exploit this by using
anthropomorphic forms in robots.
Humans seem to readily accept and prefer the 
anthropomorphic form of a robot (up to a certain point), when compared to purely mechanical 
robots. A possible explanation for this is that perceiving a system as
human-like (or also pet-like) compared to machine-like changes the 
way how users try to understand and make sense of this system. The reason for this is 
that humans tend to project everyday 
expectations of human and animalistic life onto artificial agents that use 
anthropomorphic (or zoomorphic) forms \citep{schmitz_concepts_2011}. 
Both cognitive science 
and social psychology can provide explanations for this process of projection. The basic idea 
is that \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. Studies have further 
shown that perceived similarity between the user and the robot influences the intensity of 
the participants' attributions towards the robot. Thus, the anthropomorphic design of a robot 
impacts the user's perception of the robot, and how far the user is likely to attribute and 
experience qualities such as intelligence, fun, or intentions in the robot. 
These experiences mark the transition between an immediate \textit{in-the-moment} and more thoughtful \textit{reflective} \citep{takayama_perspectives_2012} perception of, and interaction with the robot. As this distinction is also based on different cognitive processes, a more human-centered / cognitive perspective needs to be applied here, and we will treat this in Section \ref{sec:human-centered}.

In short, the central point in using anthropomorphic design in sociable robots is that 
humans tend to apply the same communication references to anthropomorphic robots 
that they would use with other humans, which in turn enhances feeling familiar and 
comfortable when interacting with the otherwise unknown system.
This positive effect has been shown in various 
empirical studies. It has been found that robots with human-like forms can enhance social 
responses from humans which in turn can positively impact on acceptance 
\citep{duffy_anthropomorphism_2003,goetz_cooperation_2002,venkatesh_theoretical_2000}. 
People responded more positively to an artifact that displayed human-like 
behavioral characteristics (emotions, facial expression) in contrast to a purely 
functional design 
\citep{eyssel_anthropomorphic_2010,krach_can_2008,reeves_media_1996,riek_how_2009}. 
Overall, these results suggest that human-like features in robots enhance 
anthropomorphism, and in turn create a social view toward the robot.

%Anthropomorphism represents just one of many examples of induction whereby "people reason about an unknown stimulus based on a better-known representation of a related stimulus" \cite{epley_when_2008}, in this case reasoning about a non-human agent based on representation of the self or other humans.

\subsubsection{Anthropomorphic form can be disadvantageous for HRI\\}

>>>>

The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like
designs of objects and non-human agents can evoke both positive and negative feelings in the human observer / user. A positive valence reflects the experience of 
emotional engagement with and feelings of empathy for the non-human agent
\citep{cheetham_human_2011}. This in turn may lead to an effective (human-like and 
human-social) interaction with the non-human agent, and increase people's acceptance of
it. 
However, at a certain point as the non-human agent
approaches but fails to closely resemble a human, the evoked valence in the
human user suddenly turns into a negative uncanny feeling of what
\cite{mori_uncanny_1970} describes as disgust and revulsion. This, in turn,
might lead to rejection of the non-human agent. The Uncanny
Valley hypothesis predicts that with further increasing the human-likeness of
the non-human agent such that it appears so realistic that it is not
distinguishable from a human but perceived to be human, the effect is again a
positive one, reflecting feelings and engagement that are similar as those
evoked by a human-human interaction. 
Of course, developers and designers of artificial agents
(\eg virtual characters, robots) try to exploit the positive effects of 
\textbf{anthropomorphic form} and avoid ``falling'' into this Uncanny Valley.
Consequently, there is a considerable research effort to
better understand the boundaries of anthropomorphic design and their effects
on the interaction and experience. To date, findings have been however
inconsistent \citep{cheetham_human_2011}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		WHAT IS ANTHROPOMORPHISM
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Towards a unified understanding: What is anthropomorphism?}
\label{sec:definitions}

Literature on anthropomorphism is quite diverse. We would like to point out that there exist different and partly conflicting definitions of the term anthropomorphism, and that in general, the \textit{terminology} is not clear. Probably as a result of this, there also exist different understandings of the \textit{phenomenon} that the term `anthropomorphism' describes.
 In this section we will first give an overview on the diversity of terms and then try to establish a common ground in the understanding of anthropomorphism (the phenomenon).

\subsection{Overview of the diversity of terms}

Originally, the term \textbf{anthropomorphism} comes from the Greek \textit{anthropos} for `man' (or `human') and \textit{morphe} for `form / structure' (or `shape'). Anthropomorphism is generally understood as people's tendency to perceive a non-human agent as if it had some human characteristics or traits. How a person perceives an agent is crucial, as it further determines how she relates to, and behaves towards the agent, or how she may behave in light of this agent \citep{epley_when_2008}.
As such, anthropomorphism can be seen as a special kind of relationship between a human interacting with a non-human agent. An anthropomorphic relationship is often characterized by the fact that the human attributes human-unique characteristics to the non-human agent. Since anthropomorphism becomes observable through these attributions, some definitions claim that anthropomorphism is not the \textit{perception} but rather the \textit{attribution} of human qualities to non-human agents. Ascribing thoughts, emotions, and intentions are given as examples, and as the most prominently anthropomorphized non-human agents animals and robots are mentioned \citep{duffy_anthropomorphism_2003,schmitz_concepts_2011}.

Apart from anthropomorphism there also exists the term \textbf{animism}, which can be defined as the attribution of life to the nonliving. The study of animism was pioneered by Piaget in the 1920s with his work on \textit{The Child's Conception of the  World}. He found that children tend to regard many inanimate objects as capable of sensations, emotions and intentions, and called this view `animism'. Since then, children`s inferences about the world have been studied intensively, and there is an agreement that children`s views and concepts of living, animal, and plant evolve as they grow up. Often synonymously used to animism is the term \textbf{zoomorphism}, which similarly applies to the cases when non-lifelike instances / artifacts are associated with animalistic attributes, excluding human-specific traits.
As `being alive' is not a human-unique property but applies to all living beings, the attribution of life may not be labeled as anthropomorphism (in a narrow understanding of it).
Another possibility to illustrate how anthropomorphism differs from attributions of animacy, is to look at paraphrases of anthropomorphism. That is to say, 
anthropomorphism can alternatively be understood as \textbf{humanization} or 
\textbf{personification} of anything other than a human being.\footnote{There exist also the phenomenons inverse to anthropomorphism, which can be found in the literature as \textbf{dehumanization} \citep{haslam_dehumanization:_2006}, or \textbf{mechanomorphism} \citep{caporael_anthropomorphism_1986}. However, we do not treat this here.} 
Robots can both be animized and anthropomorphized, but often the difference is not made.

%\hl{can be cut from here ...}
%
%Thus, animism and anthropomorphism differ in two main aspects:
%
%First, the nature of the properties which can be ascribed to an agent differs. Life-likeness is not human-unique but applies to all living beings. Thus, \textit{animism} applies to attributions of life-like properties but excludes human-specific traits. Contrary, a narrow understanding of \textit{anthropomorphism} would exclude all attributions of life-likeness that are not human-unique, such as `being alive'.  
%  
%Second, the range of agents to which life-likeness can be attributed differs. Naturally, only non-living agents can be animized. Thus, whereas non-human but \textit{living} beings (most importantly animals) can be anthropomorphized, they cannot be animized. Similarly, humans cannot neither be animized nor anthropomorphized as they are living beings and humans.
%
%\hl{... to here}

%Personification has 
%ancient roots, \eg in mythology but also in storytelling. For instance, anthropomorphism was also 
%used to describe and be able to predict natural forces or anticipate great spirits.

\subsection{Overview of the diverse understandings of anthropomorphism}

Naturally, conflicts in how terms are used lead to a diversity of understandings of the phenomenon which is described.
One reason for this might be that anthropomorphism
is a phenomenon that is studied in very different domains that might each
integrate their own understanding, or simply use the same term to refer to different things.
The conflict when reading about anthropomorphism is, that despite the fact that there is no commonly accepted definition, the terms
\textit{anthropomorphism}, \textit{anthropomorphic} or \textit{human-like} are
often used as if their meanings were clear and agreed upon
\citep{persson_anthropomorphism_2000}. It is even
argued that these terms might be misused \citep{duffy_anthropomorphism_2002,epley_when_2008}.
For instance, some researchers refer to \textit{``the
robot's level of anthropomorphism''} \cite{bartneck_is_2007}, whereas others
disagree to such a statement because they understand that a system or artifact itself does not \textit{``contain anthropomorphism''} but only gives rise to the process of anthropomorphizing in a
given user and situation. Consequently, \cite{persson_anthropomorphism_2000} conclude that
anthropomorphism emerges in the \textit{interaction} between the technology and
the user, and has to be understood as such \hl{(to which we will come back in the next section)}.

Another paradox is that, while human's tendency to anthropomorphize is commonly observed, the phenomenon is 
at the same time still rather poorly understood
\cite{epley_seeing_2007}. The phenomenon itself has been found to be very complex though sometimes
subtle and as such hard to study \cite{duffy_anthropomorphism_2002,epley_when_2008}. Further, it is not clear how to operationalize and measure anthropomorphism. We would like to discuss the challenges of studying and measuring anthropomorphism toward the end of this article, in Section \ref{sec:measuring}.

%Anthropomorphism is used in different senses throughout different disciplines 
%\citep{duffy_anthropomorphism_2003} which makes it difficult to synthesize findings. 
Depending on which definition of anthropomorphism is applied, not everything 
colloquially labeled as \emph{anthropomorphism} really is anthropomorphism.
One practical example is that, it is not always clear whether a person (may it be an adult or 
a child) who gives and anthropomorphic response, believes that the other agent really 
thinks or acts like a human or whether she is using anthropomorphism in a metaphoric 
sense \citep{leeds_childrens_1992}.
Thus, we need to be clear about what we mean when talking about anthropomorphism.

Several attempts have been made to clarify what anthropomorphism is (the phenomenon).
From a psychology point of view \cite{epley_when_2008} try to clarify things by explaining what anthropomorphism is \textit{not}. For instance, in their understanding, anthropomorphism does not include behavioral descriptions of observable actions but it requires going beyond what is directly observable. It is further proposed that anthropomorphism does not merely entail \textit{animism}, as animate life is not a \textit{uniquely} human-like characteristic.\footnote{For the whole list and more details, the reader may refer to the original work \citep{epley_when_2008}.}

%The authors list four effects that anthropomorphism does not include:
%(1) First, anthropomorphism does not include behavioral descriptions of observable actions. Anthropomorphism requires going beyond what is directly observable.
%(2) Second, anthropomorphism does not merely entail animism, as animate life is not a uniquely human-like characteristic.
%(3) Third, anthropomorphism does not include any requirement of reasoned or reflective endorsement of an inference.\footnote{\hl{(the strength of anthropomorphic inferences can vary)}}
%(4) Finally, anthropomorphism is not necessarily inaccurate. A psychological theory should take this (in)accuracy in anthropomorphism into account and predict variability in the tendency to perceive human-like traits in non-human agents.
%% comment Julia: I commented this out because I think this is too complex and too much of detail here.

As human-likeness itself is a huge dimension, some researchers argue that there might not just be one single type of anthropomorphism but different levels or gradations. For instance, \cite{persson_anthropomorphism_2000} proposes to understand anthropomorphism as a multi-layered phenomenon, and suggests that it exists of different levels, namely primitive categorization, primitive psychology, folk-psychology, traits, social roles, and emotional anthropomorphism. Correspondingly, \cite{ruijten_introducing_2014} propose to understand anthropomorphism as a continuum ranging from low to high human likeness, and to measure it as such (see Section \ref{sec:measuring}).

\subsection{Terminology and understandings used in this article}

@@

However, the term anthropomorphism should \textit{not} be used to refer to the design of a non-human agent or artifact. This would violate the understanding of anthropomorphism arising in an interaction. To refer to an imitation of human-like form in the \textit{form} of artificial agents, we propose to use the term  \textbf{anthropomorphic design}. We found that in literature, this distinction is not always done, which is a major source of confusion.

In the following, we might use the terms \textit{form} and \textit{design} interchangeably. Using the term \textit{form} to describe the design of an artifact, we refer 
not only to the shape (how it looks like) but to the total 
expression of the artifact \citep{bartneck_shaping_2004}. \cite{disalvo_hug:_2003} 
argues that \textit{form} includes a product's physical shape, materials, and 
behavioral qualities. The behavioral aspect is of special importance for interactive 
systems that show some degree of autonomy (independence).
Several categories of anthropomorphic form can be classified, but there is no 
common agreement on the borders and transitions between these categories: 
anthropomorphic, zoomorphic, caricatured, functional \citep{fong_survey_2003}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		WHY DO HUMAN ANTHROPOMORPHIZE
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Anthropomorphism is an interesting phenomenon because it tells us something about 
the relationship between the human who anthropomorphizes, and the non-human agent 
which is anthropomorphized. Though commonly observed, it still appears controversial, at 
the same time, that humans tend to \textit{``see human in the non-
human''} \citep{epley_seeing_2007}.

As anthropomorphism arises in an interaction between a human and a non-human agent
\citep{persson_anthropomorphism_2000}, one can also
take two main \textit{perspectives} when trying to explain why humans anthropomorphize \citep{lee_human_2005}: an artifact-centered perspective and a human-centered perspective (see grey triangles in Figure \ref{fig:anthropofig}). It has to be kept in mind that both perspectives are inseparable; they apply at the 
same time, and we would like to present them here just as two different viewpoints 
on the same phenomenon.

In the following we will only briefly present the artifact-centered perspective to explain anthropomorphism. Then, we provide a more detailed review and synthesis of literature that applies a human-centered perspective to explain anthropomorphism. Concretely, we integrate theories from developmental psychology, cognitive science / neuroscience, and social psychology to shed light on the question why humans anthropomorphize.

\paragraph{Artifact-centered perspective on anthropomorphism}

The artifact-centered perspective explains anthropomorphism from the ascertainable 
parts of the non-human agent, namely the anthropomorphic form of it. This research deals with the question how far anthropomorphism can be explained from the design of the artifact. Consequently, it is studied how humans perceive and respond to different types of human-like designs of non-human agents.
One of the most prominent working hypothesis in this research branch is the Uncanny Valley hypothesis \citep{mori_uncanny_1970}. In general, it is proposed that human-like designs of robots can evoke both positive and negative feelings in a human user.

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize it. 
This perspective assumes that humans directly (mindlessly) respond to life-like and social 
cues that the non-human agent or system emits \citep{nass_machines_2000}. Without thoughtful mental processing, 
humans tend to simply apply stereotypes and heuristics to the non-human agent, and in turn apply human-human social schemas and norms to the occurring interactions.
\cite{takayama_perspectives_2012} propose to refer to this immediate (sometimes visceral) sense in a situation as \textit{in-the-moment}, compared to a more distanced cogitation and consideration which she calls \textit{reflective}. As this distinction implies different cognitive processes, we will come back to it later, in section \ref{sec:cognitive-expl}.

%\subsection{An artifact-centered perspective on anthropomorphism}
%\label{sec:anthropomorphic-design}
%
%An artifact-centered perspective 
%
%The use of nature-inspired and in particular anthropomorphic (human-like) form in the 
%design of non-human things is as old and manifold as people's tendency to anthropomorphize. 
%Anthropomorphic (and life-like) forms have a long-standing tradition in design in general, and in the design of daily life things, in particular 
%\citep{norman_design_1988,guthrie_bottles_2000}. Interestingly, from its very
%beginning robots were resembling a human being in their appearance. Starting
%from when the term \textit{robot} first appeared in a play by Karel \^{C}apek in the 1920ies, robots were man-made beings, created as the \textit{``shadow of man''} \citep{replicant_????}.

%The view of people mindlessly reacting in a human-social manner to anthropomorphic forms
%is based on the fact that humans naturally react to specific 
%cues. From early childhood 
%on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
%or to respond to social cues, such as facial expressions, social gestures or turn-taking. Humans have a propensity toward particular cues that seem to trigger automatic interpersonal social responses \citep{takayama_perspectives_2012}. 
%One has to keep in mind that in 
%human-human interaction these cues are of various senses and modalities, 
%and consequently human-like design is also realized addressing several senses. 
%One of the most outstanding aspects is probably the outer appearance of an agent or 
%artifact (in particular when it is unknown or unfamiliar). \cite{schmitz_concepts_2011} 
%argues that an agent's outer appearance is likely to have an important impact on how 
%the object is overall perceived. Consequently, if a robot's outer form resembles a 
%human, the robot is likely to be treated similar to a human. Besides the physical shape of a robot, other more subtle anthropomorphic forms can be found in robots, including human-social behavioral cues (\eg eye-gaze), social gestures (\eg waving), or verbal politeness (\eg greeting, apologizing).
%People's natural reaction to these and other anthropomorphic design cues of
%robots (in particular, social or also called \textit{sociable} robots 
%\citep{breazeal_sociable_2000}) are extensively studied. This research is of particular interest in human-robot interaction as the consequences of anthropomorphic design can be both of positive and negative nature, and it is yet not clear what exactly it is that makes the difference.
%\subsubsection{How people perceive and interact with human-like robots\\}
%\label{sec:anthropomorphism-robotics}  
%Most work that studies how far human users treat robots as social agents, builds on the 
%famous work of \cite{reeves_media_1996} in HCI. In the so-called \textit{Media 
%Equation}, the authors describe that people mindlessly react in a social way to their computer 
%and that they tend to even attribute human (social) characteristics to it -- despite 
%the fact that a computer does not have human-like characteristics. Apart from 
%computers, anthropomorphism has also been observed with digital / virtual characters 
%\citep{von_der_putten_it_2010}, and artificial embodied agents, 
%such as robots \citep{rosenthal-vonderputten_experimental_2013}. 
%In short, one can say that with robots, and human-like robots in 
%particular, anthropomorphism is a commonly observed phenomenon. 
%But what are the motivations which underlie the efforts of enhancing anthropomorphism 
%through specific anthropomorphic design in robots? Why are anthropomorphic forms so 
%distinct and at the same time controversial in robotics? There is no one single answer 
%to these questions. 
%The underlying central point about the human-like design of robots lies in the 
%implications that this design holds, and the suggestions for interactions that
%it makes to the human user. 
%In other words, the appearance 
%and function of a product, whether it is an appliance or a humanoid robot, affects how 
%people perceive it, interact with it, and build long-term relationships with 
%it \citep{bartneck_shaping_2004}. 
%The principle of human-like form is probably most prominent in the design of robots, 
%due to several reasons, such as to foster efficient interaction and increase 
%acceptance. An anthropomorphic design of a robot encourages people to reason about this 
%unfamiliar system \emph{as if} it was a human. To increase this tendency, robots are 
%designed with anthropomorphic (or lifelike) forms, or are able to display and use 
%human-social cues when interacting with humans.
%However, anthropomorphic robot designs have been controversially discussed and received 
%remarkable attention since the early 1970ies, related to Mori's
%theory of the \textit{Uncanny Valley} \citep{mori_uncanny_1970}.
%%see Figure \ref{fig:uncanny_valley}.
%%\begin{figure}\centering
%%% Use the relevant command to insert your figure file.
%%% For example, with the graphicx package use
%%  \includegraphics[scale=0.75]{uncanny-valley.jpg}
%%% figure caption is below the figure
%% \caption{Mashiro Mori, illustration of the \textit{Uncanny Valley} hypothesis. The curve models the user's perceived affinity with the artifact, as its human likeness is increased. Up to a certain (not yet defined) point, a person perceives more affinity in a more human-like agent; however with an abrupt shift towards a negative perception as soon as the agent surpasses this point, of resembling a human too closely but not fully. Originally from 1970; picture from a manuscript version of \cite{mori_uncanny_2012}}
%% \label{fig:uncanny_valley}       % Give a unique label
%% \end{figure}
%Despite the fact that the Uncanny Valley theory was more a working hypothesis and not 
%meant to be empirically examined \citep{cheetham_human_2011}, it has been studied quite 
%extensively. However, results are inconsistent. We try to give a short overview (not 
%exclusive) on work related to HRI and robotics in the following. 
%\subsubsection{Anthropomorphic form can be advantageous for HRI\\}
%On one hand, developers and designers have discovered that human-like robots may better 
%match the affordances that this to humans adapted world brings along. 
%Apart from this, it is suggested that an increased human-likeness of a robot (up to a 
%certain threshold) will generally lead to 
%more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970}.
%% (see Figure \ref{fig:uncanny_valley}).
%This further increases anthropomorphism, and as such
%appears to be beneficial for human-robot relations, especially with socially 
%interactive robots \citep{fong_survey_2003}. Consequently, the domain of (social) 
%robotics tries to exploit this by using
%anthropomorphic forms in robots.
%Humans seem to readily accept and prefer the 
%anthropomorphic form of a robot (up to a certain point), when compared to purely mechanical 
%robots. A possible explanation for this is that perceiving a system as
%human-like (or also pet-like) compared to machine-like changes the 
%way how users try to understand and make sense of this system. The reason for this is 
%that humans tend to project everyday 
%expectations of human and animalistic life onto artificial agents that use 
%anthropomorphic (or zoomorphic) forms \citep{schmitz_concepts_2011}. 
%Both cognitive science 
%and social psychology can provide explanations for this process of projection. The basic idea 
%is that \textit{"the more similar in 
%appearance, the more people are likely to use themselves as a source of induction 
%and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. Studies have further 
%shown that perceived similarity between the user and the robot influences the intensity of 
%the participants' attributions towards the robot. Thus, the anthropomorphic design of a robot 
%impacts the user's perception of the robot, and how far the user is likely to attribute and 
%experience qualities such as intelligence, fun, or intentions in the robot. 
%These experiences mark the transition between an immediate \textit{in-the-moment} and more thoughtful \textit{reflective} \citep{takayama_perspectives_2012} perception of, and interaction with the robot. As this distinction is also based on different cognitive processes, a more human-centered / cognitive perspective needs to be applied here, and we will treat this in Section \ref{sec:human-centered}.
%In short, the central point in using anthropomorphic design in sociable robots is that 
%humans tend to apply the same communication references to anthropomorphic robots 
%that they would use with other humans, which in turn enhances feeling familiar and 
%comfortable when interacting with the otherwise unknown system.
%This positive effect has been shown in various 
%empirical studies. It has been found that robots with human-like forms can enhance social 
%responses from humans which in turn can positively impact on acceptance 
%\citep{duffy_anthropomorphism_2003,goetz_cooperation_2002,venkatesh_theoretical_2000}. 
%People responded more positively to an artifact that displayed human-like 
%behavioral characteristics (emotions, facial expression) in contrast to a purely 
%functional design 
%\citep{eyssel_anthropomorphic_2010,krach_can_2008,reeves_media_1996,riek_how_2009}. 
%Overall, these results suggest that human-like features in robots enhance 
%anthropomorphism, and in turn create a social view toward the robot.
%%Anthropomorphism represents just one of many examples of induction whereby "people reason about an unknown stimulus based on a better-known representation of a related stimulus" \cite{epley_when_2008}, in this case reasoning about a non-human agent based on representation of the self or other humans.
%\subsubsection{Anthropomorphic form can be disadvantageous for HRI\\}
%The anthropomorphic form also 
%increases people's expectations of the agent, and consequently, a very human-like 
%appearing robot is not only perceived but also expected to act like a human. However, today's artificial systems are 
%not able (yet) to mimic well human movements, or human social cues in an acceptable way. 
%The case is that the artificial agent pretends to be human but fails, and this is 
%more than just disappointing for the 
%human -- it appears uncanny and evokes reluctance, the so-called 
%\textit{"uncanny valley"}. The idea of naming this shift and drop from a positive to a negative reaction,
%follows Freud's description of the so-called \textit{uncanny} (a translation from the German word 
%\textit{``unheimlich''}). Controversially, something uncanny \textit{"derives its terror not from 
%something externally alien or unknown but -- on the contrary -- from something 
%strangely familiar which defeats our efforts to separate ourselves from it"} (v. Foerster, cited after \cite{hegel_understanding_2008}).
